Secure Composition of Robust and Optimising Compilers

### 2023-07-09 popl/main.tex
remove spurious text
Add todo

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[POPL '24]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

@@

  Secure compilers emit code that is hardened against adversarial target--level attacks, such as memory corruption or alteration of control--flow.
  It is an open question wether secure compilers compose in a well--behaved manner.
  This paper answers this question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
  This article discusses different variants of compositions of secure compilers in a formal manner and instantiates parts of these results in a case--study.
  To this end, an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time is presented.
  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
  The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
  The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.

@@

\begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
  $
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
                                \end{array}\right.\right\}
  $
  A trace $\trace$ is spatial memory safe ($\trace\in\smssafe$) iff given $\ev{Alloc\ \loc\ m;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ n;\sandboxtag';\securitytag'}$, then $n<m$.
\end{definition}

@@

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $
  A trace $\trace$ is memory safe ($\trace\in\mssafe$) iff $\trace\in\tmssafe$ and $\trace\in\smssafe$.
\end{definition}

@@

This section presents the key meta-theoretic results of this paper and builds on prior work~\cite{kruse2022csc}.
A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
This is not intuitive in the sense that there are a number of examples in the security domain where composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
For starters, the sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is as expected, i.e., given an $\src{L}$ program $\src{p}$, obtain its compiled $\obj{L}$ program by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

>>>>

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[POPL '24]{51st ACM SIGPLAN Symposium on Principles of Programming Languages}{January 17-19,
  2024}{London, UK}
%
%  Uncomment \acmBooktitle if th title of the proceedings is different
%  from ``Proceedings of ...''!
%
%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%  June 03--05, 2018, Woodstock, NY} 
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

@@

  Secure compilers emit code that is hardened against adversarial target--level attacks, such as memory corruption or alteration of control--flow.
  It is an open question wether secure compilers compose in a well--behaved manner.
  This paper answers this question positively: The composition of compilers that are secure with respect to some security--properties yields the intersection of these properties.
  This article discusses different variants of compositions of secure compilers in a formal manner and instantiates parts of these results in a case--study.
  To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
  Memory safety is achieved by composing secure compilers for both temporal and spatial memory safety, where the compiler for spatial memory safety is an instrumentation that inserts bounds--checks in front of pointer dereferences.
  The compiler for cryptographic constant time exploits the use of a data (operand) timing independent mode and the employed optimisations of the compilation chain are constant folding and dead code elimination.
  The compositionality results enable modular proofs for the whole compilation chain that is secure with respect to both memory safety and cryptographic constant time.

@@

\begin{definition}[\glsfirst{sms}]\label{def:trace:smsdef}
  $
  \smssafe:=\left\{\trace \left|\begin{array}{rcl}
                                  \text{If }\ev{Alloc\ \loc\ n;\sandboxtag;\securitytag}\le_{\trace}\ev{Use\ \loc\ m;\sandboxtag;\securitytag'}, \text{ then }m<n
                                \end{array}\right.\right\}
  $
\end{definition}

@@

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $
\end{definition}

@@

This section presents the key meta-theoretic results of this paper and builds on prior work~\cite{kruse2022csc}\MK{mention this at all? may reveal identity even though its double-blind}.
A main result is that secure compilers in the robust preservation framework~\cite{abate2019jour} compose.
This is not intuitive in the sense that there are a number of examples in the security domain where composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
For starters, the sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is as expected, i.e., given an $\src{L}$ program $\src{p}$, obtain its compiled $\obj{L}$ program by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
